Computing astrophysical inferences with diffusive nested sampling

### 2014-03-23 ms.tex
notation fixes

Bayesian inference is usually performed by
constructing a set of possible answers to the question of
interest, that is, the ``hypothesis space'', indexed by parameters $\pars$.
A state of prior knowledge about the parameters is described by assigning a
prior distribution $p(\pars)$. Some relevant data $\data$ is about to become
known, and prior knowledge about a connection between $\theta$ and $\data$ is
expressed by a choice of ``sampling distribution'' $p(\data | \theta)$, which
is really a family of probability distributions, one for each possible value of
$\theta$. By the product rule, the joint (prior) distribution is
\begin{eqnarray}
p(\pars, \data) = p(\pars) p(\data | \pars) = p(\data)p(\pars | \data)
\end{eqnarray}
When the
data $\data$ becomes known, our state of knowledge about $\pars$ changes
from the marginal distribution $p(\pars)$ to the conditional distribution
$p(\pars | \data_{\rm actual})$
that corresponds to the observed data. The distribution
$p(\pars | \data_{\rm actual})$ is
called the ``posterior''.
Incidentally, Bayesian inference can be understood as a special case of
MaxEnt updating, starting from the joint prior $p(\pars, \data)$ and applying
the constraint $P(\data = \data_{\rm actual}) = 1$ \citep{caticha}. Note that
this is not a ``derivation'' of Bayesian updating as the rules of probability
already imply a commitment to update from $p(\theta)$ to $p(\theta | \data)$
once $\data$ becomes known. For convenience, once the data set has been
established and attention turns to computation, the prior is denoted by
$\pi(\theta)$ and the likelihood function is $L(\theta)$.

@@

\section{Nested sampling}
The key idea of Nested Sampling \citep{skilling} is that Bayesian computation
should not solely be aimed at obtaining samples from the posterior distribution.
By beginning from the prior $\pi(\pars)$ and moving through a sequence of
distributions that successively compress the prior volume by a constant factor
that is (approximately) known, Nested Sampling can calculate posterior samples
as well as the marginal likelihood or evidence
$Z = \int p(\pars)p(\data = \data_{\rm actual} | \pars) \, d\pars$.

>>>>

Bayesian inference is usually performed by
constructing a set of possible answers to the question of
interest, that is, the ``hypothesis space'', indexed by parameters $\pars$.
A state of prior knowledge about the parameters is described by assigning a
prior distribution $p(\pars)$. Some relevant data $\data$ is about to become
known, and prior knowledge about a connection between $\theta$ and $\data$ is
expressed by a choice of ``sampling distribution'' $p(\data | \theta)$, which
is really a family of probability distributions, one for each possible value of
$\theta$. By the product rule, the joint (prior) distribution is
\begin{eqnarray}
p(\pars, \data) = p(\pars) p(\data | \pars) = p(\data)p(\pars | \data)
\end{eqnarray}
When the
data $\data$ becomes known, our state of knowledge about $\pars$ changes
from the marginal distribution $p(\pars)$ to the conditional distribution
$p(\pars | \data_{\rm actual})$
that corresponds to the observed data. The distribution
$p(\pars | \data_{\rm actual})$ is
called the ``posterior''.
Incidentally, Bayesian inference can be understood as a special case of
MaxEnt updating, starting from the joint prior $p(\pars, \data)$ and applying
the constraint $P(\data = \data_{\rm actual}) = 1$ \citep{caticha}. Note that
this is not a ``derivation'' of Bayesian updating as the rules of probability
already imply a commitment to update from $p(\pars)$ to $p(\pars | \data)$
once $\data$ becomes known. For convenience, once the data set has been
established and attention turns to computation, the prior is denoted by
$\pi(\pars)$ and the likelihood function is $L(\pars)$.

@@

\section{Nested sampling}
The key idea of Nested Sampling \citep{skilling} is that Bayesian computation
should not solely be aimed at obtaining samples from the posterior distribution.
By beginning from the prior $\pi(\pars)$ and moving through a sequence of
distributions that successively compress the prior volume by a constant factor
that is (approximately) known, Nested Sampling can calculate posterior samples
as well as the ``marginal likelihood'' or ``evidence''
$\mathcal{Z} = \int \pi(\pars)L(\pars) \, d^n\pars$.
